### Constant

To see why this approach works, recall that lines can be represented by equations.

The equation **$y = 1$** would be represented like this:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=rep(1,length(x)))
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x,x=0)) +
  theme_bw()
```

Because y is the same value (i.e., 1) at each value of x, we say that y is equal to a **"constant"**. The value of y determines the **height** of the line.


### Linear

The equation **$y = 0.5(x)$** would be represented as follows:


```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=0.5*x)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x,x=0)) +
  theme_bw()
```

Each value of y on the line is half that of x.

- When x is 2, y is 1

- When x is 5, y is 2.5

We say that the relationship between y and x is **linear**, because it is a described by a straight line, and y is simply equal to a value of x multiplied by a constant (0.5).

Multiplying x by larger numbers would make this line steeper; smaller numbers make this line less
steep. If the number were negative, the line would slope from the upper left of the plot to the lower right.


### Constant + linear component

The equation **$y = 1 + 0.5(x)$** would be represented as follows:


```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=1+0.5*x,a=rep(1,length(x)),b=0.5*x)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x,x=0)) +
  geom_line(aes(y=a),linetype=2,size=1,colour='orange') +
  geom_line(aes(y=b),linetype=2,size=1,colour='orange') +
  theme_bw()
```

We can think of this line as being made up of a constant and a linear component.

- The constant in this equation is indicated by the dashed horizontal line.
- The linear component to this equation 0.5(x) is indicated by the dashed slope line.
- The solid orange line is a combination of these two components.


### Quadratic

The equation **$y = x^2$** would be represented as follows:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=x^2)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*10,x=0)) +
  theme_bw()
```

To get each value of y, we square the value of x. 

So, when x = -5, y is 25. And if x = -4, y = 16, and so on...


**$y = -x^2$**,  would look as follows:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=-(x^2))
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*10,x=0)) +
  theme_bw()
```


### Linear plus quadratic components

The equation **$y = 50 + 5(x) - x^2$** has 

- a constant 50
- a linear component 5(x)
- a quadratic component $-x^2$:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=50+5*x-x^2,a=rep(50,length(x)),b1=5*x,b2=-(x^2))
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*10,x=0)) +
  geom_line(aes(y=a),linetype=2,size=1,colour='orange') +
  geom_line(aes(y=b1),linetype=2,size=1,colour='orange') +
  geom_line(aes(y=b2),linetype=2,size=1,colour='orange') +
  theme_bw()
```

The dashed lines on the plot indicate the intercept, linear component, and quadratic components of the equation. The solid line represents the equation.

When we see any curve, it is possible to think of it as being composed of components like this.


:::{.exercise}

Let's plot some curves. First, load the `tidyverse` package:

```{r}
library(tidyverse)
```

We can use `ggplot()` to plot curves:

```{r}
tibble(x=1:10) %>% 
  ggplot(aes(x)) + 
  stat_function(fun=function(x) 2*x + x^2)
```

`stat_function(fun=function(x) ....` allows us to superimpose any function we specify on a plot, in this case $2x + x^2$.

'x^2' means 'x-to-the-power-of-2', or $x^2$.

**Modify the equation in `stat_function(fun=function(x) EQUATION)` to plot the following:**

- $y = 3$
- $y = x$
- $y = 3 + x$
- $y = x^2$

Hint: the first one would be `stat_function(fun=function(x) 3)`

**Try to reproduce the following plot: **

(Hint: what do you have to do to the values of x to get the values of y?)

```{r,echo=F}
tibble(x=1:10) %>% 
  ggplot(aes(x)) + 
  stat_function(fun=function(x) - x^2)
```

`r hide("Answer")`

```{r,visible=F}
tibble(x=1:10) %>% 
  ggplot(aes(x)) + 
  stat_function(fun=function(x) - x^2)
```

`r unhide()`


:::


## Identifying components

It is worth noting some general characteristics of polynomial equations to give you a better idea of what these various components in the equation look like.

### Linear (first-order) components

In a linear equation of the form **$y = b_0 + b_1(x)$**, the coefficient $b_0$ represents the **intercept**. $b_1$ is the coefficient for x.

The line for a linear equation keeps on increasing and decreasing for all values of x, with no bends:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=x)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x,x=0)) +
  theme_minimal()
```


### Quadratic (second-order) components

Equations with a quadratic component have form **$y = b_0 + b_1(x) + b_2(x^2)$**. Again, $b_0$ represents the intercept. $b_1$ and $b_2$ are the coefficients for $x$ and $x^2$, respectively.

The **maximum/minimum** of a curve is the **point of inflection** at which the curve switches from decreasing to increasing (or increasing to decreasing). Equations with a quadratic component have one minimum (if the $x^2$ component is positive) or one maximum (if the $x^2$ component is negative).

For curves with positive $x^2$ components:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=50+5*x+x^2)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*20,x=0)) +
  geom_point(y=min(d$y),x=-2.5,size=4) + 
  theme_bw()
```

The dot indicates the minimum. Either side of this point, the values of y increase.



For curves with $-x^2$ components:

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=50+5*x-x^2)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*10,x=0)) +
  geom_point(y=max(d$y),x=2.5,size=4) +
  theme_bw()
```

The dot indicates the maximum. Either side of this point, the values of y decrease.

:::{.exercise}

Use `ggplot()` and `stat_function()` to plot a 2nd degree (quadratic) polynomial with the following characteristics:

- a negative intercept
- a positive linear component
- a negative quadratic component


`r hide("Show me an example")`
```{r}
tibble(x=1:10) %>% 
   ggplot(aes(x)) + 
   stat_function(fun=function(x) -5 + 0.5*x - 6.0*x^2) +
   xlim(-5,5) + 
   ylim(-100,100)
```

`xlim()` is used to specify the range of values on the x-axis (-5 to 5).

`ylim()` is used to specify the range of values on the y-axis (-100 to 100).

The curve specified in the line `stat_function(fun=function(x) -5 + 0.5*x - 6.0*x^2)` is 

$y = -5 + 0.5x - 6.0x^2$

:::

### Cubic (third) order components

Curves with a cubic component have _two minima/maxima_ (i.e., two points of inflection), and are of the form **$y = b_0 + b_1(x) + b_2(x^2) + b_3(x^3)$**, where $b_0$ is the intercept, and $b_1$, $b_2$, and $b_3$ are the coefficients for the $x$, $x^2$, and $x^3$ components. 

```{r,echo=F}
d<-tibble(x=seq(-10,10,0.1),y=0 + 0.2*x + 0.2*x^2 + 0.2*x^3)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*30,x=0)) +
  theme_bw()
```

:::{.exercise}

Use `ggplot()` and `stat_function()` to plot a 3rd degree (cubic) polynomial with the following characteristics:

- a positive intercept
- a negative linear component
- a positive quadratic component
- a negative cubic component


`r hide("Show me an example")`
```{r}
tibble(x=1:10) %>% 
   ggplot(aes(x)) + 
   stat_function(fun=function(x) 5 - 0.5*x + 6.0*x^2 - 0.5*x^3) +
   xlim(-10,10) + 
   ylim(0,1000)
```

The curve specified in the line `stat_function(fun=function(x) 5 - 0.5*x + 6.0*x^2 - 0.5*x^3)` is

$y = 5 - 0.5x + 6.0x^2 - 0.5x^3$

:::

### Quartic (fourth) order components

A curve with a quartic component would have _three maxima/minima_ (i.e., three points of inflection). The equation would be of the form **$y = b_0 + b_1(x) + b_2(x^2) + b_3(x^3) + b_4(x^4)$**.

```{r,echo=F,warning=FALSE}
d<-tibble(x=seq(-10,10,0.1),y=0+0.5*x-6.0*x^2+1.0*x^3+0.5*x^4)
d %>% ggplot(aes(x=x)) + 
  geom_line(aes(y=y),size=2,colour='orange') +
  geom_line(aes(y=0)) +
  geom_line(aes(y=x*30,x=0)) +
  ylim(-100,100) +
  theme_bw()
```

:::{.exercise}

Use `ggplot` and `stat_function` to plot a 4th degree polynomial.

Note: The points of inflection may not always be visible if the scale on either the y- or x-axis is very large. You can use `ylim()` and `xlim()` to adjust the scale of the axes, e.g., for the quartic function above:

`r hide("Show me")`
```{r}
tibble(x=1:10) %>% 
   ggplot(aes(x)) + 
   stat_function(fun=function(x) 0 + 0.5*x - 6.0*x^2 + 1.0*x^3 + 0.5*x^4) +
   xlim(-5,5) + 
   ylim(-100,100)
```
The code specified in `stat_function(fun=function(x) 0 + 0.5*x - 6.0*x^2 + 1.0*x^3 + 0.5*x^4)` is

$y = 0 + 0.5x - 6.0x^2 + 1.0x^3 + 0.5x^4)$

`r unhide()`


:::
