```{r, include=F}
library(tidyverse)
library(webex)
```

<!--

## Glossary


### Aesthetics and `aes` in ggplot

Aesthetics means how things look. In ggplot the `aes` command creates a mapping between variables in
a dataset and the visual features (aesthetics) of the plot

 -->

## Cheatsheet {#cheatsheet}

This covers everything we have done to-date (or will soon), in abbreviated form. You will need to
load tidyverse for most of these examples:

```{r}
library(tidyverse)
```

### Basics

Assign a value to a name (variable):

```{r}
meaningoflife <- 43
```

Simple arithmetic

```{r}
2+2
2*2
2^3 # 2 cubed
```

Compare values:

```{r}
2 == 2  # note the doubled = sign
2 != 4  # 2 is not equal to 4
2 > 2
2 >= 2
```

### Sequences

```{r}
1:10
seq(1,10)
seq(1,100, by=7) # 7 times table
```

For strings use `c()`

```{r}
c("Wibble", "Wobble", "Nobble")
```

Combinations of sequences:

```{r}
expand.grid(
   colour=c("Red", "Green"),
   position=c("Top", "Bottom"))
```

### Randomness

A random sample:

```{r}
sample(1:20)
```

Random-normal numbers:

```{r}
rnorm(n=20, mean=0, sd=5)
rnorm(20, 0, 5)  # equivalent to line above(but this is a different sample!)
```

Other types of random numbers:

```{r}
runif(10)
rbinom(10, size=1, prob=.5) # like a coin toss where 1=heads
rbinom(10, size=5, prob=.5) # like best of 5 coin tosses
```

### Loading data

CSV:

```{r, eval=F}
df <- read_csv('filename.csv') # from a file on your server account
df <- read_csv('<URL>') # direct from the web
```

SPSS:

```{r , eval=F}
library(haven)  # load haven first
df_from_spss <- read_spss('filename.sav')
```

Excel (note different sheets are loaded from same file):

```{r, eval=F}
expt1 <- read_excel('simple-excel-example.xlsx', sheet="Experiment 1")
expt2 <- read_excel('simple-excel-example.xlsx', sheet="Experiment 2")
```

### 'Looking at' datasets

If a dataset is large we don't want to look at it all at once.

Show the first 3 rows for all variables:

```{r}
iris %>% head(3)
```

Show a list of all the **_columns_** with as many datapoints as fit in the window:

```{r}
iris %>% glimpse
```

Calculate some useful summaries of all variables in the dataset:

```{r}
library(skimr)
iris %>% skim()
```

Check for patterns of [missing data](#missing-data):

```{r message=F}
library(mice)
boys %>% md.pattern()
```

### Choosing columns and rows {#dplyr-cheat}

#### Selecting columns

```{r}
mtcars %>% select(mpg, wt) %>% head()
```

Selecting columns which start with a particular string:

```{r}
iris %>% select(starts_with('Sepal')) %>% head
```

Selecting columns which match a name:

```{r}
iris %>% select(matches('Width')) %>% head
```

#### Renaming columns

```{r}
mtcars %>% select(wt) %>% rename(weight=wt) %>% head
```

:::{.tip}

Try to avoid spaces or punctuation in your variable names if possible.

:::

If you do end up with spaces in your column names, you can still access them by putting 'backticks'
around the name.

```{r include=F}
annoying_dataframe = tibble(`What is your favourite colour?` = c("Red", "Blue", "Green" ))
```

Our annoying dataframe is like this:

```{r}
annoying_dataframe  %>% head
```

Rename the column:

```{r}
annoying_dataframe %>%
  rename('favourite_colour' = `What is your favourite colour?`) %>%
  head
```

#### Filtering rows

Select rows where a variable matches a particular number:

```{r}
gapminder::gapminder %>% filter(year==2002) %>% head
```

Select rows where a variable matches a particular string value (letters/words):

```{r}
iris %>% filter(Species=="setosa") %>% head
```

Select rows meeting (both of) 2 criteria:

```{r}
mtcars %>% filter(wt > 3 & cyl == 4)
```

Select rows meeting either one criteria OR another:

```{r}
mtcars %>% filter(wt > 3 | cyl == 4) %>% head
```

### Groups and summaries {#cheat-group-summary}

Calculate statistics on a dataset:

```{r}
iris %>% summarise(mean(Sepal.Length), mean(Sepal.Width))
```

Give our summary variables specific names (these are bad names, don't do this!):

```{r}
iris %>% summarise(wibble=mean(Sepal.Length), wobble=mean(Sepal.Width))
```

Calculate summaries for groups in the a dataset:

```{r}
iris %>% group_by(Species) %>% summarise(mean(Sepal.Length), sd(Sepal.Length))
```

### Plotting

Also see the [section on visualisation](#visualisation).

Density plot with colours:

```{r}
diamonds  %>%
  ggplot(aes(x=price, y=..scaled.., color=factor(color))) +
  geom_density()
```

Scatterplot, with facets:

```{r}
diamonds %>%
  ggplot(aes(x=carat, y=price)) +
  geom_point() +
  facet_wrap(~color)
```

Boxplot, with added line:

```{r}
diamonds %>%
  ggplot(aes(x=cut, y=carat)) +
  geom_boxplot() +
  geom_hline(yintercept=.5, color="red", linetype="dotted")
```

Scatterplot with logarithmic scale on x axis, with point sizes scaled (to country-size `pop`):

```{r}
gapminder::gapminder %>%
  filter(year==2002) %>%
  ggplot(aes(gdpPercap, lifeExp, size=pop)) +
  geom_point() +
  scale_x_log10()
```
