# FAQ Data Analysis and Visualisation Task {#FAQ-assessment}


```{r, echo=F, include=F}
knitr::opts_chunk$set(echo = TRUE, collapse=TRUE, cache=TRUE, comment=">", message=FALSE)
library(tidyverse)
library(webex)
library(pander)
theme_set(theme_minimal())
```


\

---

### How do I add a caption to a summary output table?

`pander()` can be used to output a summary table with a caption. Use the `caption=“XXX”` argument. For example:


```{r}
library(tidyverse)
library(pander)

mtcars %>% 
  group_by(cyl) %>% 
  summarise(mean_mpg = mean(mpg)) %>% 
  pander(caption = "Mean miles per gallon by cyl")

```


\

---


\

### What does it mean if RStudio prints out "e-" next to a number? {#e-meaning}

If R prints "1.4e-4", in the output this actually means "$1.4 \times 10^{-4}$", or "0.00014". It is a way of printing out very small (or very large) numbers.

"e-4" means "move the decimal point 4 places to the left". See this by executing the command below -- it should return "0.00014".


```{r}
1.4e-4
```

So:

- "e-5" means move the decimal point 5 places to the left (e.g., `2.5e-5` is "$2.5 \times 10^{-5}$", or 0.000025)
- "e-10" means move the decimal point 10 places to the left (e.g., `2.5e-10` is "$2.5 \times 10^{-10}$", or 0.00000000025).
- "e5" means move the decimal point 5 places to the **right** (e.g., `2.5e5` is "$2.5 \times 10^{5}$", 250000).


\

---

\

### How do I set up an R Markdown file?

R-markdown is a way of embedding R code, figures, and text in a single executable document. (As an example, all the worksheets for this module were written in r-markdown.)

If you want to create an R markdown file, go to File > New File > R Markdown.

Slides to help get started are here: https://chrisjberry.github.io/datafluencyCB/slides/PSYC753_Chris2_Rmd.pptx

There's a good cheat sheet here:
- [R Markdown](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)

<!--





### How can I round numbers to X decimal places? 

The `round()` function can be used to round a number. For example, the command:

```{r}
round(3.14159, 2)
```

rounds 3.14159 to 2 decimal places.

We can combine `round` with commands from the `dplyr` package to round **all** of the numbers in a data frame to two decimal places in one go. 

First look at the built-in dataset `DNase`:

```{r}
DNase %>% head()
```

The columns `conc` and `density` contain numeric values that we may want to display to 2 decimal places.

We can achieve that using `mutate_if()` from the `dplyr` package, as follows:

```{r,eval=F}
DNase %>% mutate_if(is.numeric, round, 2) 
```


**Explanation**: `mutate_if` will change the entry in the data frame **if** some condition is satisfied. In this case, if the entry is a number (i.e.,  `is.numeric`), it will apply `round` to it to `2` decimal places. (The column `Run` is not rounded because it is is an ordinal, rather than a numeric, variable.)

<div class='exercise'>
Try rounding the values in `DNase` to one decimal place.

`r hide("Show me")`
```{r,eval=FALSE}
DNase %>% mutate_if(is.numeric, round, 1) 
```
`r unhide()`

</div>



\

---

### Cheat sheets

Some handy cheat sheets for RStudio are located here: https://www.rstudio.com/resources/cheatsheets/

The following may be of particular use for topics from this module:

- [Data transformation with dplyr](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
- [Data visualisation with ggplot](https://github.com/rstudio/cheatsheets/raw/master/data-visualization-2.1.pdf)
- [R Markdown](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf)


\

---






\

---

### How can I save a file from RStudio to my own hard drive/computer so I can submit it to the DLE?

If using R Studio through the university website (recommended) [rstudio.plymouth.ac.uk](https://rstudio.plymouth.ac.uk), you can save a file in the Files pane as follows:

- Tick the check box next to the file you want to save.
- Click the "More" button > Export. 
- You can then choose where to save the file on your computer.


-->

